Leadership across cultures and contexts: an exploratory analysis

### 2020-04-09 Response to reviewers.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Response to editor and reviewer comments

### Editor comments

**Statistical reporting and further analysis: The GLMM should be reported in full somewhere. This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers? What about the national origin effects for the ethnographers? We recommend taking seriously the reflexive analytical techniques of putting the ethnographer in the analysis (see work by Robert Aunger working in the Congo Basin). Does the ethnographer matter? If so, how much?**

Although we determined the gender of almost all ethnographers (see below), we were not able to confirm the national origins of most.

We now briefly describe the random effects for ethnographers. [need to do]

**Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population? This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**

We determined the gender of almost all authors of all documents included in our study. We found that of the 884 text records with no female author or co-author, 21 (2.4%) reported female leadership. Of the 225 text records with a female author or co-author, 20 (8.9%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders. After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = 3.5), it was not statistically significant (p = 0.25). We now report this in our results section.

@@

## Reviewer 1 comments

**This manuscript presents a descriptive analysis of the HRAF to provide information about leadership in traditional societies. Overall, I enjoyed reading the paper and thought it was a useful resource. It provides a comprehensive analysis of leadership across a wide swath of traditional cultures. It reports on the types of functions leaders serve, the qualities possessed by leaders, and the costs and benefits of leadership experienced by both leaders and followers. The paper will likely be of interest to anyone working in the area of leadership, social hierarchy, or group behavior. I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**

We greatly appreciate the detailed comments and very useful suggestions.

**The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines. Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies. The paper appears to miss out on a useful opportunity to provide some commentary on this issue. Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**

We agree our initial submisson did not sufficently accomplish this goal. We have now added summarizing, concluding statements on the qualities and functions of leaders in WEIRD and non-WEIRD settings in the introduction.

### 2020-04-09 leadership_across_cultures_contexts.Rmd
Resolving merge conflicts across 2 .R and 2 .Rmd files. No issues.

Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven. We draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967]. The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers. For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive. See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.

The operationalized qualities and functions (See Table \@ref(tab:qualfuncvars)) were developed from the ethnographic text records in the database. First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars). ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phratries, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).

@@

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture. The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(leader_text2)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

>>>>

Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven. We draw on the same database of `r nrow(leader_text2)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967]. The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethographic coverage from each of 60 culture regions. It comprises over `r trunc(sum(dfwc['N Pages eHRAF'])/1000)`K pages of digitized text.

The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers. For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive. See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.

The operationalized qualities and functions (See Table \@ref(tab:qualfuncvars)) were developed from the ethnographic text records in the database. First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. Then, KS coded all `r nrow(leader_text2)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars). ZG then followed the same process of coding the entire sample of `r nrow(leader_text2)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).

@@

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture. Our search yielded `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures. The number of text records per culture ranged from `r min(dfwc$number_leader_records, na.rm=T)` to `r max(dfwc$number_leader_records, na.rm=T)`, with a median of `r median(dfwc$number_leader_records, na.rm=T)`, a mean of `r signif(mean(dfwc$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(dfwc$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(leader_text2)` text records, `r final_record_count` had evidence for at least one variable in the study.

```{r corpus}
cor_pages_tr <- cor.test(~number_leader_records + `N Pages eHRAF`, dfwc)
```

There was a weak positive relationship between the number of leader text records we retrieved from each culture and the total number of pages of ethnography for each culture in eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`). A loess regression suggested that the number of records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure \@ref(fig:eHRAFpages)).

@@

# Supplementary information

```{r eHRAFpages, fig.height=6, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF. Each dot is one culture. Line fit by loess regression."}
plot_pages_tr
```

We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` variables: publication date, female vs. male-only authorship, and variation in the total number of pages of ethnography for each culture. For each binary study variable, we fit a logistic mixed effects model with that variable as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures. We fit two sets of models. The first univariate set fit separate models for each study variable and each bias variable (total number of models = 105x3 = 315). The second multivariate set included all bias variables in each model (total number of models = 105). All p-values were corrected for a FDR of 5%. After correction, the coefficients of female authorship and pages of ethnography were not significant in any model. Seven variables had significant associations with year of publication in both the univariate or multivariate sets (with nearly identical coefficients), but Group determination/cohesiveness was only significant in the univariate model, and Leader social service benefits was only significant in the multivariate model. See Figure \@ref(fig:pubdate).

```{r pubdate, fig.cap="The effect of publication date on the probability of finding evidence for leader qualities, functions, benefits, and costs. Multivariate values are  controlling for female authorship and number of pages of ethonography for each culture. All models had random intercepts for ethnographers nested within cultures. Models were fit for each of the 105 variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant. Publication date was centered and standardized by 1 SD. Each dot is one coefficient (converted from log odds to odds ratios). Bars are 95% CI."}
plot_pubdate
```

## Feature models

```{r features, fig.width=10, fig.cap="Marginal effects of Poission mixed effects regression models of Feature variables by mode of subsistence. Only models whose subsistence coefficient had a corrected p-value < 0.05. Base level: agriculturalists. See text for details."}

x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))

```
